### 2024-03-09 archived/table9.tex
Initial Overleaf Import

                &\multicolumn{5}{c}{}                                      \\
                &     MENA&      SSA&Diff. (MENA - SSA)         &     S.E.&     Obs.\\
\midrule
\hspace{0.25cm} Age 65+ (years)&    5.440&    3.383&    2.057\sym{***}&  (0.045)&    32272\\
\addlinespace
\hspace{0.25cm} GDP per capita (USD)&30289.436& 4790.638&25498.798\sym{***}&(290.617)&    31787\\
\addlinespace
\hspace{0.25cm} Life expectancy (years)&   75.657&   62.719&   12.938\sym{***}&  (0.053)&    32751\\
\addlinespace
\hspace{0.25cm} Extreme poverty (\%)&    4.289&   37.920&  -33.630\sym{***}&  (0.194)&    22908\\
\addlinespace
\hspace{0.25cm} Population density (per sq km)&  298.756&  108.561&  190.195\sym{***}&  (5.288)&    31807\\