### 2024-03-09 manuscript.tex
Updates from Overleaf

% Reference TeX file for: https://medium.com/the-stata-guide/the-stata-to-latex-guide-6e7ed5622856
% Asjad Naqvi (asjadnaqvi@gmail.com)
% Last update: 25 September, 2022

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                    %%%
%%%      Preamble      %%%
%%%                    %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{article}

%%% packages
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{amssymb, amsfonts, amsmath}
\usepackage{bm}
\usepackage[margin=1in]{geometry} % full-width
    \topskip        =   20pt
    \parskip        =   10pt
    \parindent      =   0 pt
    \baselineskip   =   15pt

\usepackage{setspace}               % line spacing
    \onehalfspacing
\usepackage{adjustbox}              % resizing
\usepackage{caption}                % to reset the cations etc 
\usepackage{rotating}               % for the sidewaystable
    \numberwithin{table}{section}   % reset the Table numbering for each section
\usepackage{booktabs}               % neatly formatting lines
\usepackage{dcolumn}                % aligning decimals
    \newcolumntype{d}[1]{D{.}{.}{#1}}
\usepackage{natbib}                 % bibliography
    \bibliographystyle{plainnat}

\newcommand{\sym}[1]{\rlap{#1}} % for the stars

%\usepackage{threeparttable}    
%\usepackage{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                    %%%
%%%     Front page     %%%
%%%                    %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Stata--to--\LaTeX Guide\thanks{See the article on \href{https://medium.com/the-stata-guide/the-stata-to-latex-guide-6e7ed5622856}{Medium} for details.}}

\author{
    Asjad Naqvi\thanks{\href{mailto:asjadnaqvi@gmail.com}{asjadnaqvi@gmail.com}. For comments and suggestions, please open an \href{https://github.com/asjadnaqvi/The-Stata-Guide/issues}{Issue} on the \href{https://github.com/asjadnaqvi/The-Stata-Guide}{Stata Guide's GitHub} page. }
    }

\date{Last updated: 8 June, 2023}

\begin{document}
\maketitle

\begin{abstract}
This article introduces various wrappers for the Stata \texttt{estout} package \citep{Jann2005,Jann2007} that allows one to generate various summary statistic and regression tables in \LaTeX. Rather than going into the details of the \texttt{estout} package options, which are fairly extensive, the aim of this document is to provide ``templates'' for easy  replication. This is an evolving document so please check regularly for updates. Since \LaTeX requirements change over time, please see the change log for details, or message me to fix errors if they occur. And please feel free to request or suggest other table formats. 
\end{abstract}

\vspace{4ex}
\small{\textit{Keywords}: Stata, \LaTeX, estout, tables, regressions}

\small{\textit{JEL}: 1337}

%\begin{center}
%\textbf{DOCUMENT IS BEING LIVE UPDATED! USE WITH CAUTION.}    
%\end{center}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                    %%%
%%%   Introduction     %%%
%%%                    %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

\listoftables

\newpage
\section*{Introduction}
The document is divided into two sections. 
Section \ref{sec:equations} is a simple guide for writing equations in \LaTeX. It highlights some basic mistakes users name.

\vspace{5pt}

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                   %%%
%%%    Regressions    %%%
%%%                   %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\section{Regressions} \label{sec:regressions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Regression 1: Basics  %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{regression1_1.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                      %%%
%%%   Bibliography       %%%
%%%                      %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

%\bibliographystyle{apalike}
\bibliography{bibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%    END OF DOCUMENT    %%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% more ideas for tables? Just email or msg on Twitter